Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2022-02-24 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex

A GP model with a periodic covariance function does not fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation. In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic square exponential covariance function.

@@

Next, we analyze how the interaction between $m$ and $c$ affects the quality of the approximation. The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models. Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and $c$. Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model. Figure \ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models. Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$. Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application. We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure~\ref{fig3_Post_part1}. From these results, some general conclusions may be drawn:

\begin{itemize}
\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
\item There exists a minimum $c$ below which an accurate approximation will never be achieved regardless of the number of basis functions $m$.
\end{itemize}

\subsection{Near linear proportionality between $m$, $c$, and $\ell$} \label{subsec_theoretic_linearity}

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_relationships.png}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
  \label{fig6_relationships}
\end{figure*}

Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ relative to the half-range $S$ of the input domain). Figure \ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension. More precisely, for a given GP model (with a particular covariance function) with length-scale $\frac{\ell}{S}$ (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq. (\ref{eq_diff_covs}). We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq. (\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
%
\begin{equation} \label{eq:rel_total_distance}
 \int | k(\tau) - \tilde{k}_m(\tau)|  \,\mathrm{d}\tau  < 0.01 \int k(\tau) \,\mathrm{d}\tau,
\end{equation}
%
where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions. Alternatively, these figures can be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$. Of course, we may also read it as providing the minimum $\ell$ that can be approximated with high accuracy given $m$ and $c$. We obtain the following main conclusions:

\begin{itemize}
\item As $\ell$ increases, $m$ required for an accurate approximation decrease.
\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
\item For a given $\ell$ there exist a minimum $c$ under which an accurate approximation is never going to be achieved regardless of $m$. This fact can be seen in Figure~\ref{fig6_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$). As $\ell$ increases, the minimum valid $c$ also increases.
\end{itemize}

@@

\paragraph*{Mat\'ern ($\nu$=3/2)}:
%
\begin{align}
&m= 3.42 \, \frac{c}{\nicefrac{\ell}{S}} \;\; \Leftrightarrow \;\; \nicefrac{\ell}{S}= 3.42 \, \frac{c}{m}, \label{eq_m_l_mat32}\\
&\text{with} \nonumber \\
&c \, \geq \, 4.5 \, \nicefrac{\ell}{S} \;\; \& \;\, c \geq \, 1.2 \label{eq_c_vs_l_mat32}
\end{align}
%
These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \rightarrow \infty$). Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions. If the aim is to find minimal $m$ to speedup the computation, a further refined formula can be obtained for new covariance functions.

Figure~\ref{fig6_relationships} and previous equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$. Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$. More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions. For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$. Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases. However, as an approximation, we can use the unidimensional GP conclusions in Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.

@@

As commented in Section~\ref{sec_method_periodic}, in Appendix~\ref{sec_periodic} we present an approximate linear representation of a periodic squared exponential covariance function. In Appendix~\ref{sec_periodic}, we also analyze the accuracy of this linear representation and derive the minimum number of terms $m$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel. Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition. This relationship between $m$ and $\ell$ for a periodic squared exponential covariance function is gathered in Figure~\ref{figB1_m_lscale_periodic} and the numerical equation was estimated in eq.~\eqref{eq_J_l_periodic} which is depicted next:
%
\begin{align*}
&m \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{m}
\end{align*}

@@

As stated above, equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used) provide the minimum length-scale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data. Basically, if the length-scale estimate is accurate, we can expect the HSGP approximation to be accurate as well.

@@

\item[A1.] Make an initial guess on the lengthscale $\ell_1$. If there is no useful information available, we recommend to start with a large lengthscale, about $\ell_1 \in [0.5, 1]$. Set the iteration index to $i=1$.

\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}.

\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_lscale_comparison.png}}
\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP lengthscale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
  \label{fig7_lscale_comparison}
\end{figure*}

In this section, we first illustrate that accurate estimates of the length-scale implies accurate approximations via HSGPs. Figure~\ref{fig7_lscale_comparison} left shows a comparison of the length-scale estimates
obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness. Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq. (\ref{eq_c_vs_l_QE}). Figure \ref{fig7_lscale_comparison} right shows the RMSE of the HSGP models with the exact GP model as the reference. It can be seen that accurate estimates of the length-scale imply small RMSEs.

@@

\citet{solin2014explicit} derive a cosine series expansion for the periodic covariance function (\ref{eq_cov_periodic}) as follows,
% 
\begin{equation} \label{eq_cov_periodic_taylor_approx}
k(\tau)= \alpha \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 \tau),
\end{equation}
which comes basically from a Taylor series representation of the periodic covariance function. The coefficients $\tilde{q}_j^2$ 
%
\begin{equation} \label{eq_q}
\tilde{q}_j^2= \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
\end{equation}
where $j=1,2,\cdots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
%
\begin{equation} \label{eq_q0}
\tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
\end{equation}
The covariance in eq.~(\ref{eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation. This approximation converges to eq.~(\ref{eq_cov_periodic}) when $J \rightarrow \infty$.

An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{eq_q}) and (\ref{eq_q0}), and thus leading to the following variance coefficients:
%
\begin{equation}\label{eq_q_2}
\begin{split}
\tilde{q}_j^2=& \frac{2\mathrm{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
\tilde{q}_0^2=& \frac{\mathrm{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
\end{split}
\end{equation} 
for $j=1,2,\cdots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind. This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required. A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function are provided by \citet{solin2014explicit}.

>>>>

A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation. In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic squared exponential covariance function.

@@

Next, we analyze how the interaction between $m$ and $c$ affects the quality of the approximation. The length-scale and marginal variance of the covariance function will no longer be fixed but instead we compute the joint posterior distribution of the function values and the hyperparameters using the dynamic HMC algorithm implemented in Stan \citep{StanTeam:2021} for both the exact GP and the HSGP models. Figure~\ref{fig3_Post_part1} shows the posterior predictive mean and standard deviation of the function as well as the covariance function obtained after fitting the model for varying $m$ and~$c$. Figure~\ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the exact GP model. Figure~\ref{fig5_lscale_vs_J} shows the estimated length-scale and marginal variance for the exact GP model and the HSGP models. Looking at the RMSEs in Figure~\ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computation time for this example would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$. Further, the less conservative choice of $m = 10$ and $c = 1.5$ could also produce a sufficiently accurately approximation depending on the application. We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure~\ref{fig3_Post_part1}. From these results, some general conclusions may be drawn:
%
\begin{itemize}
\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq.~\eqref{eq_eigenvalue}, where $L$ appears in the denominator.
\item There exists a minimum $c$ below which an accurate approximation will never be achieved regardless of the number of basis functions $m$.
\end{itemize}

\subsection{Near linear proportionality between $m$, $c$ and $\ell$} \label{subsec_theoretic_linearity}

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_relationships.png}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the squared exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
  \label{fig6_relationships}
\end{figure*}

Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $\frac{\ell}{S}$ (lengthscale $\ell$ relative to the half-range $S$ of the input domain). Figure \ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with squared exponential, Mat\'ern ($\nu$=3/2), and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension. More precisely, for a given GP model (with a particular covariance function) with length-scale $\frac{\ell}{S}$ (From now to later on we will refer to $\frac{\ell}{S}$ by just $\ell$) and given a boundary factor $c$, Figure \ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq. (\ref{eq_diff_covs}). We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq. (\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
%
\begin{equation} \label{eq:rel_total_distance}
 \int | k(\tau) - \tilde{k}_m(\tau)|  \,\mathrm{d}\tau  < 0.01 \int k(\tau) \,\mathrm{d}\tau,
\end{equation}
%
where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions. Alternatively, these figures can be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$. Of course, we may also read it as providing the minimum $\ell$ that can be approximated with high accuracy given $m$ and $c$. We obtain the following main conclusions:
%
\begin{itemize}
\item As $\ell$ increases, $m$ required for an accurate approximation decrease.
\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
\item For a given $\ell$ there exist a minimum $c$ under which an accurate approximation is never going to be achieved regardless of $m$. This fact can be seen in Figure~\ref{fig6_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$). As $\ell$ increases, the minimum valid $c$ also increases.
\end{itemize}

@@

\paragraph*{Mat\'ern ($\nu$=3/2)}:
%
\begin{align}
&m= 3.42 \, \frac{c}{\nicefrac{\ell}{S}} \;\; \Leftrightarrow \;\; \nicefrac{\ell}{S}= 3.42 \, \frac{c}{m}, \label{eq_m_l_mat32}\\
&\text{with} \nonumber \\
&c \, \geq \, 4.5 \, \nicefrac{\ell}{S} \;\; \& \;\, c \geq \, 1.2 \label{eq_c_vs_l_mat32}
\end{align}
%
These constants vary monotonically with respect to $\nu$ (squared exponential corresponding to Mat\'ern with $\nu \to \infty$). Using the formula for Mat\'ern ($\nu$=3/2) provides the largest $m$ and $c$, and thus this formula alone could be used as a conservative choice for all Mat\'ern covariance functions with $\nu \geq 3/2$ and likely as a good initial guess for many other covariance functions. If the aim is to find minimal $m$ to speedup the computation, a further refined formula can be obtained for new covariance functions.

Figure~\ref{fig6_relationships} and previous equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$. Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$. More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions. For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$. Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases. However, as an approximation, we can use the unidimensional GP conclusions in Figure~\ref{fig6_relationships} or equations~(\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.

@@

As commented in Section~\ref{sec_method_periodic}, in Appendix~\ref{sec_periodic} we present an approximate linear representation of a periodic squared exponential covariance function. In Appendix~\ref{sec_periodic}, we also analyze the accuracy of this linear representation and derive the minimum number of terms $m$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel. Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition. This relationship between $m$ and $\ell$ for a periodic squared exponential covariance function is gathered in Figure~\ref{figB1_m_lscale_periodic} and the numerical equation was estimated in eq.~\eqref{eq_J_l_periodic} which is depicted next:
%
\begin{align*}
&m \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{m} .
\end{align*}

@@

As stated above, equations \eqref{eq_m_l_QE}, \eqref{eq_m_l_mat52}, \eqref{eq_m_l_mat32} and \eqref{eq_J_l_periodic} (depending on which kernel is used) provide the minimum length-scale that can be closely accurately given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of estimating the latent function from the data. Basically, if the length-scale estimate is accurate, we can expect the HSGP approximation to be accurate as well.

@@

\item[A1.] Make an initial guess on the length-scale $\ell_1$. If there is no useful information available, we recommend to start with a large length-scale, about $\ell_1 \in [0.5, 1]$. Set the iteration index to $i=1$.

\item[A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section~\ref{sec_num_equations}.

\item[A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section~\ref{sec_num_equations}.

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_lscale_comparison.png}}
\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP length-scale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
  \label{fig7_lscale_comparison}
\end{figure*}

In this section, we first illustrate that accurate estimates of the length-scale implies accurate approximations via HSGPs. Figure~\ref{fig7_lscale_comparison} left shows a comparison of the length-scale estimates
obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness. Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq.~(\ref{eq_c_vs_l_QE}). Figure~\ref{fig7_lscale_comparison} (right) shows the RMSE of the HSGP models with the exact GP model as the reference. It can be seen that accurate estimates of the length-scale imply small RMSEs.

@@

\citet{solin2014explicit} derive a cosine series expansion for the periodic covariance function (\ref{eq_cov_periodic}) as follows,
% 
\begin{equation} \label{eq_cov_periodic_taylor_approx}
k(\tau)= \alpha \sum_{j=0}^{J} \tilde{q}_j^2 \cos(j\omega_0 \tau),
\end{equation}
which comes basically from a Taylor series representation of the periodic covariance function. The coefficients $\tilde{q}_j^2$ 
%
\begin{equation} \label{eq_q}
\tilde{q}_j^2= \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!},
\end{equation}
where $j=1,2,\ldots,J$, and $\lfloor \cdot \rfloor$ denotes the floor round-off operator. For the index $j=0$, the coefficient is 
%
\begin{equation} \label{eq_q0}
\tilde{q}_0^2= \frac{1}{2} \frac{2}{\exp(\frac{1}{\ell^2})} \sum_{j=0}^{\lfloor \frac{J-j}{2} \rfloor} \frac{(2\ell^2)^{-j-2}}{(j+i)!i!}.
\end{equation}
The covariance in eq.~(\ref{eq_cov_periodic_taylor_approx}) is a $J$th order truncation of a Taylor series representation. This approximation converges to eq.~(\ref{eq_cov_periodic}) when $J \rightarrow \infty$.

An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{eq_q}) and (\ref{eq_q0}), and thus leading to the following variance coefficients:
%
\begin{equation}\label{eq_q_2}
\begin{split}
\tilde{q}_j^2=& \frac{2\mathrm{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
\tilde{q}_0^2=& \frac{\mathrm{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
\end{split}
\end{equation} 
for $j=1,2,\ldots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind. This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required. A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function are provided by \citet{solin2014explicit}.
